Cryptanalysis of SBC based post-quantum signature scheme \\\textit{Graduation internship report}

### 2025-07-02 rapport.tex
fin partie sur SBC

		\subsection{Multiparty computation protocol for the SBC problem}
		A Multi-Party Computation (MPC) protocol is an interactive protocol between $N$ parties $\mathcal{P}_1, \dots, \mathcal{P}_N$, who wish to jointly compute a function $f$ over their private inputs $x^{\lir}$. At the end of the protocol, each party outputs its own computed value, denoted $f_i(x) = f(x^{\lir})$.

@@

		\begin{definition}[def 2.4.5 \cite{feneu2023}]
			Let $N$ and $t$ be integers such that $1 \leq t \leq N$. Let $\Pi_f$ be an MPC protocol with $N$ parties $\mathcal{P}_1,\dots,\mathcal{P}_N$, computing a function $f$. The protocol $\Pi_f$ is $t$-private in the semi-honest model if for all $I \subset [N]$ such that $|I| \leq t$, there exists a $PPT$ algorithm $\mathcal{S}$ such that $\mathcal{S}(I, \llbracket x \rrbracket_I, f_I(x))$ is perfectly indistinguishable from the joint distribution of the views of the parties in $I$, where $f_I(x) := {f_i(x), \forall x \in I}$.	
		\end{definition}

@@

		\begin{itemize}
			\item \textbf{Addition}: each party $\mathcal{P}_i$ can locally compute $\llbracket x + y \rrbracket$ by adding their shares:
			$$
				(x + y)^{\lir} = x^{\lir} + y^{\lir} \text{ } \forall i \in [N].
			$$
			We denote this by $\llbracket x + y \rrbracket = \llbracket x \rrbracket + \llbracket y \rrbracket$.
			\item \textbf{Adding a constant}: we can easily share the constant $\alpha$ among the parties as $\llbracket \alpha \rrbracket = (\alpha, 0, \dots, 0)$.
			The parties can then compute $\llbracket x + \alpha \rrbracket = \llbracket x \rrbracket + \llbracket \alpha \rrbracket$. We denote this by $\llbracket x \rrbracket + \alpha$.
			\item \textbf{Multiplication by a constant}: each party $\mathcal{P}_i$ can locally compute $\llbracket \alpha x \rrbracket$ by multiplying their respective share by $\alpha$:
			$$
				(\alpha x)^{\lir} = \alpha x^{\lir} \text{ } \forall i \in [N]. 
			$$
			We denote this by $\llbracket \alpha x \rrbracket = \alpha \llbracket x \rrbracket$.
		\end{itemize}

@@

		A malicious party $\mathcal{P}_i$ could exploit this by waiting to receive all the shares of $\llbracket t_0 \rrbracket$ from the other parties, and then choosing their own share $t_0^{\lir}$ such that
		$$
			t_0 = \sum_{j \neq i} t_0^{\llbracket j \rrbracket} + t_0^{\lir}
		$$
		is a root of the polynomial. This would allow $\mathcal{P}_i$ to bias the evaluation towards a value that satisfies the equation, which could compromise the protocol by causing it to output \textit{accept} even though $(\vec{x}, \vec{y}) \notin \text{NSBC}(\vec{u}, \vec{v})$.

		To mitigate this issue, parties will not be allowed to choose their share of $\llbracket t_0 \rrbracket$ arbitrarily. Instead, each share $t_0^{\lir}$ will be deterministically generated as the output of a hash function applied to a fixed, party-specific string concatenated with a common seed. Specifically, the share is computed as:
		$$
		t_0^{\lir} = \mathcal{H}(\text{"}t_0\text{ value for party }i\text{"} \, \| \, \text{seed}),
		$$
		where $\mathcal{H}$ is a cryptographic hash function and $\|$ denotes concatenation. This ensures that the shares are unpredictable yet deterministic, thereby preventing any party from biasing the outcome.

@@

		\subsection{Identification Protocol}
		We present the identification protocol, which serves as a basis for constructing the signature scheme using the Fiat-Shamir transformation.

		\subsection{Keys and signature sizes}
		The size of the private key is just the size of the seed that is used to generate it which is derived from a PRF. So we only need to store the seed which is of size $\lambda$. So the size of the private key is $16$ bytes in practice. For the public key, we saw that all of it is generated randomly using a PRF and $v_n$ is computed. The size of $v_n$ is $2\lambda$ so the is size of public key is $3\lambda$ which in practice is $48$ bytes.

@@

		\begin{lemma}\label{UB-KOA}
			Let $(q,n,k)$ be a set of parameters for the NSBC problem. Assume that the NSBC function $\text{NSBC} :\mathbb{F}_{q^k}^{2n} \rightarrow \mathbb{F}_{q}^{2n}$ is $(\tau, \epsilon_{OWF})$-hard, meaning that no probabilistic polynomial-time ($\mathcal{PPT}$) adversary $mathbb{A}$ running in time at most $\tau$ can solve the NSBC problem with probability greater than $\epsilon_{OWF}$.
			\begin{align*}
				\text{Succ}(\mathcal{A}) = \mathcal{P}\left[
				\begin{array}{l}
					(u, v) \leftarrow \mathbb{F}_{q^k}^{2n}, \\
					(x, y) \leftarrow \mathcal{A}(u, v), \\
					(x, y) \in \text{SBC}[u, v]
				\end{array}
				\right] < \epsilon_{OWF}.
			\end{align*}
			Then the signature scheme is \textbf{UB-KOA}.
		\end{lemma}
		The proof of \ref{UB-KOA} can be found in \ref{appendix:proof}.

		\begin{theorem}\label{EUF-CMA}
			Assume that no adversary can distinguish the output of the pseudorandom function (PRF) used in the protocol from truly random bits. The adversary outputs a bit $b$, where $b=0$ indicates that they believe the bits originate from the PRF, and $b=1$ indicates they believe the bits are uniformly random
			$$
				| \mathcal{P}_{seed \in \{0,1\}^n}[b=0 \mid p \leftarrow PRF(seed)] - \mathcal{P}[b=1 \mid p \overset{{\scriptscriptstyle\$}}{\leftarrow} \{0,1\}^{2n}] | \leq \frac{1}{2}.
			$$
			Suppose also that the protocol uses a cryptographic hash function which is one-way and does not have collision attacks with better complexity than $2^{\lambda}$ using a $\mathcal{PPT}$, where $\lambda$ is the number security bits needed for the signature scheme.

>>>>

		\subsection{Multiparty computation protocol for the SBC problem}
		A Multi-Party Computation (MPC) protocol is an interactive protocol between $N$ parties $P_1, \dots, P_N$, who wish to jointly compute a function $f$ over their private inputs $x^{\lir}$. At the end of the protocol, each party outputs its own computed value, denoted $f_i(x) = f(x^{\lir})$.

@@

		\begin{definition}[def 2.4.5 \cite{feneu2023}]
			Let $N$ and $t$ be integers such that $1 \leq t \leq N$. Let $\Pi_f$ be an MPC protocol with $N$ parties $P_1,\dots,P_N$, computing a function $f$. The protocol $\Pi_f$ is $t$-private in the semi-honest model if for all $I \subset [1, N]$ such that $|I| \leq t$, there exists a $PPT$ algorithm $\mathcal{S}$ such that $\mathcal{S}(I, \llbracket x \rrbracket_I, f_I(x))$ is perfectly indistinguishable from the joint distribution of the views of the parties in $I$, where $f_I(x) := {f_i(x), \forall x \in I}$.	
		\end{definition}

@@

		\begin{itemize}
			\item \textbf{Addition}: each party $P_i$ can locally compute $\llbracket x + y \rrbracket$ by adding their shares:
			$$
				(x + y)^{\lir} = x^{\lir} + y^{\lir} \text{ } \forall i \in [1, N].
			$$
			We denote this by $\llbracket x + y \rrbracket = \llbracket x \rrbracket + \llbracket y \rrbracket$.
			\item \textbf{Adding a constant}: we can easily share the constant $\alpha$ among the parties as $\llbracket \alpha \rrbracket = (\alpha, 0, \dots, 0)$.
			The parties can then compute $\llbracket x + \alpha \rrbracket = \llbracket x \rrbracket + \llbracket \alpha \rrbracket$. We denote this by $\llbracket x \rrbracket + \alpha$.
			\item \textbf{Multiplication by a constant}: each party $P_i$ can locally compute $\llbracket \alpha x \rrbracket$ by multiplying their respective share by $\alpha$:
			$$
				(\alpha x)^{\lir} = \alpha x^{\lir} \text{ } \forall i \in [1, N]. 
			$$
			We denote this by $\llbracket \alpha x \rrbracket = \alpha \llbracket x \rrbracket$.
		\end{itemize}

@@

		A malicious party $P_i$ could exploit this by waiting to receive all the shares of $\llbracket t_0 \rrbracket$ from the other parties, and then choosing their own share $t_0^{\lir}$ such that
		$$
			t_0 = \sum_{j \neq i} t_0^{\llbracket j \rrbracket} + t_0^{\lir}
		$$
		is a root of the polynomial. This would allow $P_i$ to bias the evaluation towards a value that satisfies the equation, which could compromise the protocol by causing it to output \textit{accept} even though $(\vec{x}, \vec{y}) \notin \text{NSBC}(\vec{u}, \vec{v})$.

		To mitigate this issue, parties will not be allowed to choose their share of $\llbracket t_0 \rrbracket$ arbitrarily. Instead, each share $t_0^{\lir}$ will be deterministically generated as the output of a hash function applied to a fixed, party-specific string concatenated with a common seed and something inherit to the party like their shares. Specifically, the share is computed as:
		$$
		t_0^{\lir} = \mathcal{H}(\text{"}t_0\text{ value for party }i\text{"} \, \| \, \text{seed} | x^{\lir}, y^{\lir}, X_1^{\lir}, X_2^{\lir}, Y_1^{\lir}, Y_2^{\lir}, A^{\lir}, B^{\lir}),
		$$
		where $\mathcal{H}$ is a cryptographic hash function and $\|$ denotes concatenation. This ensures that the shares are unpredictable yet deterministic, thereby preventing any party from biasing the outcome.

@@

		\subsection{Identification Protocol}
		We present the identification protocol, which serves as the foundation for constructing the signature scheme via the Fiat–Shamir transformation.

		To do so, we transform the MPC protocol into an interactive ZKP of knowledge of the witness $(\vec{x}, \vec{y})$ such that it is a solution of $\text{NSBC}[\vec{u}, \vec{v}]$.

		The prover $\mathcal{P}$ simulates the MPC protocol \emph{in their head} and commits to the views of each party $h = \mathcal{H}_V(View_1, \dots, View_N).$ The view of each party consists of all the messages they send and receive during the protocol. In addition, $\mathcal{P}$ commits to the following values:
		\[
		X_1 + t_0(\vec{u} \cdot \vec{x}), \quad X_1 + t_0(\vec{v} \cdot \vec{x}), \quad Y_1 + t_0(\vec{v} \cdot \vec{y}), \quad Y_2 + t_0(\vec{u} \cdot \vec{y}), \quad A + Bt_0.
		\]

		Then, the verifier $\mathcal{V}$ selects a random index $i^* \in [1, N]$. The prover $\mathcal{P}$ opens all the shares of the following values for all $i \in [1, N]$ such that $i \neq i^*$:
		\[
		x^{\lir}, \quad y^{\lir}, \quad X_1^{\lir}, \quad X_2^{\lir}, \quad Y_1^{\lir}, \quad Y_2^{\lir}, \quad A^{\lir}, \quad B^{\lir}.
		\]

		Using this information, $\mathcal{V}$ can reconstruct the shares of $t_0^{\lir}$ as previously described. Moreover, they can compute the corresponding shares of the committed values:
		\[
		(X_1 + t_0(\vec{u} \cdot \vec{x}))^{\lir}, \quad (X_1 + t_0(\vec{v} \cdot \vec{x}))^{\lir}, \quad (Y_1 + t_0(\vec{v} \cdot \vec{y}))^{\lir}, \quad (Y_2 + t_0(\vec{u} \cdot \vec{y}))^{\lir}, \quad (A + Bt_0)^{\lir}, \quad \forall i \neq i^*.
		\]

		Since the sharing is linear, the verifier can efficiently recover the missing shares corresponding to $i^*$. For the detailed description of these computations, we refer the reader to~\cite{HJ23}.

		Finally, the verifier $\mathcal{V}$ can compute $h' = \mathcal{H}_V(View_1, \dots, View_N)$ and compare with $h$. If they match, they \textit{accept} and \textit{reject} otherwise.

		We recall that, to reduce the communication cost, we use correlated GGM (cGGM) trees and transmit only the co-paths of the tree instead of all values. Additionally, the HyperCube technique is employed to reduce the computational overhead.

		To convert the identification scheme into a signature scheme, we apply the Fiat-Shamir transformation by replacing the verifier's random choice of $i*$ with the output of a hash function.

		The soundness of the protocol is given by $\frac{1}{N} + \frac{1}{q^k}$. This is because the underlying MPC protocol is $(N-1)$-private, so the prover $\mathcal{P}$ can potentially cheat on only one view, and the verifier catches them with probability $\frac{1}{N}$. Furthermore, we have shown that a false positive may occur when $t_0$ happens to be a root of a certain polynomial, which occurs with probability $\frac{1}{q^k}$.

		Rather than increasing the number of parties $N$, which would increase computational cost, we repeat the protocol $\tau$ times such that:

		$$
		\left( \frac{1}{N} + \frac{1}{q^k} \right)^{\tau} < 2^{-\lambda},
		$$
		where $\lambda$ is the targeted security level.

		\subsection{Keys and signature sizes}
		The size of the private key is just the size of the seed that is used to generate it which is derived from a PRF. So we only need to store the seed which is of size $\lambda$. So the size of the private key is $16$ bytes in practice with the proposed parameters. For the public key, we saw that all of it is generated randomly using a PRF and $v_n$ is computed. The size of $v_n$ is $2\lambda$ so the is size of public key is $3\lambda$ which in practice with the proposed parameters is $48$ bytes.

@@

		\begin{lemma}\label{UB-KOA}
			Let $(q,n,k)$ be a set of parameters for the NSBC problem. Assume that the NSBC function $\text{NSBC} :\mathbb{F}_{q^k}^{2n} \rightarrow \mathbb{F}_{q}^{2n}$ is $(\tau, \epsilon_{OWF})$-hard, meaning that no probabilistic polynomial-time ($\mathcal{PPT}$) adversary $mathbb{A}$ running in time at most $\tau$ can solve the NSBC problem with probability greater than $\epsilon_{OWF}$.
			\begin{align*}
				\text{Succ}(\mathcal{A}) = \mathbb{P}\left[
				\begin{array}{l}
					(u, v) \leftarrow \mathbb{F}_{q^k}^{2n}, \\
					(x, y) \leftarrow \mathcal{A}(u, v), \\
					(x, y) \in \text{SBC}[u, v]
				\end{array}
				\right] < \epsilon_{OWF}.
			\end{align*}
			Then the signature scheme is \textbf{UB-KOA}.
		\end{lemma}
		The proof of \ref{UB-KOA} can be found in \ref{appendix:proof}.

		\begin{theorem}\label{EUF-CMA}
			Assume that no adversary can distinguish the output of the pseudorandom function (PRF) used in the protocol from truly random bits. The adversary outputs a bit $b$, where $b=0$ indicates that they believe the bits originate from the PRF, and $b=1$ indicates they believe the bits are uniformly random
			$$
				| \mathbb{P}_{seed \in \{0,1\}^n}[b=0 \mid p \leftarrow PRF(seed)] - \mathbb{P}[b=1 \mid p \overset{{\scriptscriptstyle\$}}{\leftarrow} \{0,1\}^{2n}] | \leq \frac{1}{2}.
			$$
			Suppose also that the protocol uses a cryptographic hash function which is one-way and does not have collision attacks with better complexity than $2^{\lambda}$ using a $\mathcal{PPT}$, where $\lambda$ is the number security bits needed for the signature scheme.
